---
title: "Thrive by Five Adjustments for 2021-2024 Comparison"
author: "Tyryn Carnegie"
date: last-modified
format:
  html:
    theme: lumen
    code-fold: true
    embed-resources: true    
    highlight-style: tango
    toc: true
    toc-depth: 5
    number-sections: false
    anchor-sections: false
execute:
  warning: false
---

In the doc "Thrive by Five 2024 - Province score investigation" we found there to be assessor bias in the 2021 Thrive data. In this document we'll be incorporating findings from that document to adjust the 2021 and 2024 datasets so that they are comparable. The steps are as follows:

1. Remove assessor 2681 from the 2021 dataset, others may need to be removed too. 
2. Readjust the all outliers' scores (identified via Cohen's D) using mixed-effects model approach, using the assessor random effect coefficients to adjust domain scores. 
3. In the 2024 data, remove all ELPs with fewer than 6 children enrolled in the 4-year old age group.

After the above steps we then estimate the proportion of children On Track in 2021 and 2024. We then implement ...

### Set up


```{r}
pacman::p_load(
  dplyr,
  ggplot2,
  tidyr,
  lubridate,
  stringr,
  haven,
  scales,
  lme4,
  lmerTest,
  glmmTMB,
  flextable,
  leaflet,
  elomtools,
  skimr
)
source("functions.R")
# 2021 data --------------------------------------------------------------------
tb5_2021 <- read_dta("data/thrive_by_five_20230830.dta") #%>% # Merged with the ECD census to get missing fee values

# 2024 data --------------------------------------------------------------------
tb5_2024 <- read_dta("data/2024tb5_all_forms_inter_data.dta") #%>% 
  # Generate variables to align with functions 

# Get the id_enumerator variable from an earlier dataset
tb5_2024_id_enumerator  <- read_dta("data/tb5_2024_elom_elp_client.dta") %>% 
  select(id_child, date, id_enumerator)
tb5_2024 <- tb5_2024 %>% 
  left_join(tb5_2024_id_enumerator, by = join_by(id_child))

# Generate new fee variable
tb5_2021 <- tb5_2021 %>% 
  mutate(fees_amount = ifelse(fees==0, 0, fees_amount)) %>% 
  mutate(fees_amount_deflated =  fees_amount*1.089) 

# Generate a fee group variable according to the 2024 cuts
tb5_2021 <- tb5_2021 %>% 
  mutate(n_pri_fee_group7 = case_when(
    fees_amount_deflated<=60 ~ 1, # These thresholds were taken from the 2024 data
    fees_amount_deflated>60 & fees_amount_deflated<=120 ~ 2,
    fees_amount_deflated>120 & fees_amount_deflated<=240 ~ 3,
    fees_amount_deflated>240 & fees_amount_deflated<=350 ~ 4,
    fees_amount_deflated>350 & fees_amount_deflated<=800 ~ 5,
    fees_amount_deflated>800 & fees_amount_deflated<=1690 ~ 6,
    fees_amount_deflated>1690 ~ 7,
    TRUE ~  NA
  ))


```


# Remove problematic assessors

Here we remove assessors that clearly assessed children incorrectly in the 2021 sample. Referring to the other document the following assessors were identified:

- 2681

There are arguably others too. There are several assessors who scored exactly the same way on every assessment for certain items but weren't Cohen's D outliers. These include:

- 344
- 351
- 1732

These assessors should arguably be dropped ahead of 2681 as these assessors will not see their scores adjusted as per the Cohen's D outlier identification approach. Or we should at least also have their domain scores imputed too? 

```{r, echo=TRUE}
tb5_2021 <- tb5_2021 %>% 
  filter(id_enumerator!="2681")

``` 


# Identify problematic assessors via Cohen's D

Like with the doc "Thrive by Five 2024 - Province score investigation" I highlight assessors in the 2021 sample whose average ELOM scores sit more than 2 standard deviations from the mean. This time, however, we're using a dataset that has dropping assessor 2681. 

```{r}
#| label: fig-2021-Cohens-D
#| fig-cap: "2021 Cohen's D outliers"
#| fig-height: 10
#| fig-cap-location: top

plot_enumerator_cohens_d(tb5_2021, score_var = total_elom)

``` 


# Impute domain scores of problematic assessors


```{r}
outliers <- c("2762", "2669", " 571","2976", "2685", " 2796")
elom_long3 <- tb5_2021 %>%
  # Coerce to the same variable names as Colin's code
  select(id_child=child_id, id_enumerator,elom_domain_1=domain_1, elom_domain_2=domain_2, elom_domain_3=domain_3, elom_domain_4=domain_4, elom_domain_5=domain_5, elom_child_age=child_age, elom_child_sex=child_gender, n_prov_fee_quintile_rev=n_pri_fee_group7) %>% 
  pivot_longer(
    cols       = c(starts_with("elom_domain_"), -contains("cuts")),
    names_to   = "domain", 
    values_to  = "domain_score"
  ) 

# Fit the mixed-effects model
#    random intercepts for child, domain, enumerator.
fit <- lmer(
  domain_score ~ elom_child_age + elom_child_sex + n_prov_fee_quintile_rev +
    (1 | id_child) +         # random intercept for child
    (1 | domain) +           # random intercept for domain
    (1 | id_enumerator),     # random intercept for enumerator
  data = elom_long3
)
# summary(fit)

# Identify problematic assessors
prob_enums <- outliers # There are 7


re_enum <- ranef(fit)$id_enumerator[["(Intercept)"]]

# Convert to a data frame with enumerator ID and the random intercept
df_enum <- data.frame(
  id_enumerator = rownames(ranef(fit)$id_enumerator),
  enum_re       = re_enum
)

elom_long3 <- elom_long3 %>%
  mutate(id_enumerator = as.character(id_enumerator)) %>% 
  left_join(df_enum, by = "id_enumerator") %>%
  # if some enumerators are missing in ranef (possible if data is incomplete),
  # coalesce them to 0
  mutate(enum_re = coalesce(enum_re, 0))

elom_long3 <- elom_long3 %>%
  mutate(
    # If enumerator is in the problem list, subtract their random effect.
    # If not in problem list, do no correction to raw score.
    domain_score_adj = if_else(
      id_enumerator %in% prob_enums,
      domain_score - enum_re,   # remove enumerator effect
      domain_score             # keep as-is
    )
  )

# Threshold adjusted domain scores at 0 and 20
elom_long3 <-  
  elom_long3 %>% 
    mutate (domain_score_adj = case_when(
      domain_score_adj < 0 ~ 0,
      domain_score_adj > 20 ~ 20,
      .default = domain_score_adj
    ))

# Reshape the data
elom_adj_wide <- elom_long3 %>% 
  select(id_child, domain, child_age=elom_child_age, domain_score_adj) %>% 
  pivot_wider(id_cols=c(id_child, child_age), names_from=domain, values_from = domain_score_adj) %>% 
  mutate(score = elom_domain_1 + elom_domain_2 + elom_domain_3 + elom_domain_4 + elom_domain_5) %>% 
  rename(domain_1=elom_domain_1, domain_2=elom_domain_2, domain_3=elom_domain_3, domain_4=elom_domain_4, domain_5=elom_domain_5)

elom_adj_wide <- fn_age_group(elom_adj_wide, tool_in = "4and5")

# Regenerate total cuts
elom_adj_wide <- apply_domain_categories(elom_adj_wide, tool_in="4and5", "1")
elom_adj_wide <- apply_domain_categories(elom_adj_wide, tool_in="4and5", "2")
elom_adj_wide <- apply_domain_categories(elom_adj_wide, tool_in="4and5", "3")
elom_adj_wide <- apply_domain_categories(elom_adj_wide, tool_in="4and5", "4")
elom_adj_wide <- apply_domain_categories(elom_adj_wide, tool_in="4and5", "5")
elom_adj_wide <- apply_domain_categories(elom_adj_wide, tool_in="4and5", "score")

elom_adj_wide <- elom_adj_wide %>% 
  select(id_child, domain_1:score, domain_1_cuts:total_elom_cuts) %>% 
    mutate(across(
    ends_with("_cuts"),
    ~ factor(
        dplyr::recode(as.character(.),
          "1" = "Falling Far Behind",
          "2" = "Falling Behind",
          "3" = "On Track",
          .default = NA_character_
        ),
        levels = c("Falling Far Behind", "Falling Behind", "On Track"),
        ordered = TRUE
      )
  )) %>%
  rename_with(~ paste0(.x, "_adj"), domain_1:total_elom_cuts)


# Merge this into the 2021 data
tb5_2021 <- tb5_2021 %>% 
  left_join(elom_adj_wide, by=join_by("child_id"=="id_child"))


```


# Remove ELPs with fewer than 6 children aged 4-5 in the 2024 sample

In the 2021 Thrive by Five only ELPs with 6 or more children were included in the study. In order to make the samples more comparable, here we remove ELPs with fewer than 6 children from the 2024 sample. 

Here we use the variable count_month_total from the 2024 sampling form - "How many of the enrolled children were born between X and Y". This gives the number enrolled within the valid age range. This is probably different to how this was done in 2021 (e.g., instead asking "How many children are 4 years old?")

```{r}
# Using the variable count_month_total from the sampling form to get the size of the ELP with eligible children
sample_size <- read_dta("data/tb5_2024_sample_client.dta") %>% 
  select(id_ecd, n_children, count_children, count_month_total,count_year, n_children)


tb5_2024_filtered <- tb5_2024 %>% 
  left_join(sample_size) %>% 
  filter(count_month_total>=6)



```

Taking this step reduces the 2024 sample size substantially from `r nrow(tb5_2024)` to `r nrow(tb5_2024_filtered)`.


# Entropy balancing to mathc

```{r}


```


