---
title: "Language investigation"
author: "Tyryn Carnegie"
date: last-modified
format:
  html:
    theme: lumen
    code-fold: true
    embed-resources: true    
    highlight-style: tango
    toc: true
    toc-depth: 5
    number-sections: false
    anchor-sections: false
execute:
  warning: false
---

```{r}
pacman::p_load(
  dplyr,
  cem,
  ggplot2,
  tidyr,
  lubridate,
  stringr,
  haven,
  scales,
  lme4,
  purrr,
  lmerTest,
  glmmTMB,
  flextable,
  leaflet,
  elomtools,
  survey,
  skimr,
  WeightIt,
  cobalt
)
source("functions.R")

# 2021 data --------------------------------------------------------------------
tb5_2021 <- read_dta("data/thrive_by_five_20230830.dta") #%>% # Merged with the ECD census to get missing fee values

# 2024 data --------------------------------------------------------------------
tb5_2024 <- read_dta("data/2024tb5_all_forms_inter_data.dta") #%>% 
  # Generate variables to align with functions 

# Get the id_enumerator variable from an earlier dataset
tb5_2024_id_enumerator  <- read_dta("data/tb5_2024_elom_elp_client.dta") %>% 
  select(id_child, date, id_enumerator)
tb5_2024 <- tb5_2024 %>% 
  left_join(tb5_2024_id_enumerator, by = join_by(id_child))


``` 


# The proportion of assessments where the ELOM assessment language matches the ELOM form's child home language
The table below shows the proportion of assessment where the ELOM assessment language variable (elom_language_assessment) matches the child home language captured on the ELOM form (elom_language_child). While all provinces are either 100% or near 100%, its curious as to why they aren't all 100%. We would expect language mismatches to have been dropped. 

```{r}
tb5_2024_lang_elom <- tb5_2024 %>%
  transmute(
    elom_prov,
    assess = as.integer(zap_labels(elom_language_assessment)),
    child  = as.character(zap_labels(elom_language_child))
  ) %>%
  mutate(
    # normalize spacing and pad with spaces to make whole-number word-boundary matches easy
    child_norm = paste0(" ", str_squish(child), " "),
    language_match = if_else(
      !is.na(assess) & !is.na(child_norm) &
        str_detect(child_norm, paste0("\\b", assess, "\\b")),
      1L, 0L, missing = 0L
    )
  ) %>%
  select(elom_prov, elom_language_assessment = assess,
         elom_language_child = child, language_match)

tbl <- tb5_2024_lang_elom %>% 
  # mutate(language_match = as.integer(as.character(elom_language_assessment) ==
  #                                    as.character(cg_language_home))) %>% 
  # filter(!is.na(cg_language_home), cg_language_home != "", cg_language_home != "None") %>% 
  group_by(elom_prov) %>%
  summarise(prop_language_match = round(100 * mean(language_match), 1),
            n_language = n(), .groups = "drop") %>% 
  rename(
    Province = elom_prov,
    # `Assessment language` = elom_language_assessment,
    `% ELOM home languages matches assessment language` = prop_language_match,
    N = n_language
  ) 

tbl %>% 
  flextable()

```

# The proportion assessed where the assessment language matches the MAIN home language in the PCG form

The table below shows the % of assessed children for whom the assessment language matches the language the child mostly speaks at home (as per the PCG form, cg_language_home). 

The table shows the language outcomes for the 2024 sample only. The table's results are notable for:

- 25% of Eastern Cape assessments were conducted in English when the child mostly spoke another language at home
- Gauteng's result were striking. 39% of Gauteng assessments were conducted in English when the child mostly spoke another language at home. Similarly for Sesotho for assessments (24%), Setswana (29%), Sepedi (37%).
- 54% of Northern Cape assessments were conducted in English when the child mostly spoke another language at home. 
- 41% of North West assessments were conducted in English when the child mostly spoke another language at home. 
- 22% of Western Cape assessments were conducted in English when the child mostly spoke another language at home. 

```{r}
#| label: tbl-lang-match
#| tbl-cap: "The proportion assessed where the assessment language matches the main home language"

lang_cols <- paste0("cg_languages_", 1:11)

tb5_2024_lang <- tb5_2024 %>% 
  select(id_child, elom_prov, cg_language_home, elom_language_assessment) %>% 
  as_factor() 

tb5_2024_home_langs <- tb5_2024 %>% 
  select(id_child, all_of(lang_cols)) %>% 
  pivot_longer(
    cols = starts_with("cg_languages_"),
    names_to = "name",
    values_to = "value"
  ) %>% 
  filter(as.integer(value) == 1) %>% 
  group_by(id_child) %>% 
  filter(n() == 1) %>% 
  ungroup() %>% 
  mutate(cg_language_coerce = as.integer(str_remove(name, "^cg_languages_")),
         cg_language_coerce = factor(cg_language_coerce, levels=c(1:11),
                                     labels = c("Afrikaans", 
                                                "English",
                                                "isiNdebele",
                                                "isiXhosa",
                                                "isiZulu",
                                                "Sesotho",
                                                "Sepedi",
                                                "Setswana",
                                                "siSwati", 
                                                "Tshivenda",
                                                "Xitsonga"))) %>% 
  select(id_child, cg_language_coerce)

tb5_2024_lang <- tb5_2024_lang %>%
  left_join(tb5_2024_home_langs, by = "id_child") %>%
  mutate(cg_language_coerce = factor(cg_language_coerce,
                                     levels = levels(cg_language_home)),
         cg_language_home = if_else(is.na(cg_language_home),
                                    cg_language_coerce, cg_language_home)) %>%
  select(-cg_language_coerce) %>% 
  mutate(elom_language_assessment =
           forcats::fct_recode(elom_language_assessment,
                      "Sepedi" = "Sesotho se Leboa (Sepedi)"))


# build the summary table
tbl <- tb5_2024_lang %>% 
  mutate(language_match = as.integer(as.character(elom_language_assessment) ==
                                     as.character(cg_language_home))) %>% 
  filter(!is.na(cg_language_home), cg_language_home != "", cg_language_home != "None") %>% 
  group_by(elom_prov, elom_language_assessment) %>% 
  summarise(prop_language_match = round(100 * mean(language_match), 1),
            n_language = n(), .groups = "drop") %>% 
  rename(
    Province = elom_prov,
    `Assessment language` = elom_language_assessment,
    `% most used home language matches assessment language` = prop_language_match,
    N = n_language
  )

# color palette: low values = red, high = white 
pal <- col_numeric(
  palette = c("#b2182b", "#fddbc7", "#ffffff"), # deep red → light red → white
  domain  = c(0, 100)
)

# build flextable
ft <- flextable(tbl) %>%
  colformat_num(j = "% most used home language matches assessment language",
                digits = 1, suffix = "%")

# apply background gradient to that column
col_vals <- tbl$`% most used home language matches assessment language`
ft <- bg(
  ft,
  j = "% most used home language matches assessment language",
  i = seq_len(nrow(tbl)),
  bg = pal(col_vals)
)

# optional: improve contrast for very low values (white text on dark red)
ft <- color(
  ft,
  j = "% most used home language matches assessment language",
  i = which(col_vals <= 25),  # threshold is your call
  color = "white"
) %>%
  autofit()

ft

```

## The impact of language mismatch on ELOM outcomes
To explore this finding more, here we compare the relationship between Total ELOM scores and not having the assessment conducted in the child's home language. Below we use a simple mixed effects model where ECD centre, enumeration area and stratum are used as random effects. Fixed effects include child age, child sex, height-for-age, ELP fee level, and lastly an indicator saying whether the assessment was conducted in the most spoken language at the child's home. 

**The table below shows that when the language of assessment matches the language most spoken at home, the child on average 1.12 points more on average.** 

```{r}
library(lme4)
library(sjPlot)

tb5x <- tb5_2024 %>%
  left_join(tb5_2024_lang , by = "id_child") %>%
  as_factor() %>% 
  mutate(
    # language_match
   language_match = as.integer(as.character(elom_language_assessment.y) ==
                                     as.character(cg_language_home.y)),
    # outcome: 1 = On Track
    ontrack = as.integer(n_elom_total_cuts == "On Track"),
    # tidy covariates
    elom_child_sex = as.factor(elom_child_sex),
    n_pri_fee_group7 = as.factor(n_pri_fee_group7)
  )

mod <- lmer(
  n_elom_total_score ~ elom_child_age + elom_child_sex + n_elom_zhfa + n_pri_fee_group7 +
    language_match +
    (1 | id_ecd) + (1 | id_ea) + (1 | stratum),
  data = tb5x,
  # family = binomial(link = "logit") #,
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e5))
)


tab_model(
  mod,
  dv.labels  = "ELOM total score",
  pred.labels = c(
    "(Intercept)",
    "Child age (months)"        = "elom_child_age",
    "Girl (ref: Boy)"           = "elom_child_sexGirls",
    "Height-for-age (z)"        = "n_elom_zhfa",
    "Fee group (factor)"        = "n_pri_fee_group7",
    "Assessment language = home language" = "language_match"
  ),
  # show.ci   = TRUE,
  show.icc  = TRUE,             # intraclass correlation
  show.r2   = TRUE,             
  show.se   = TRUE,
  show.re.var = TRUE            # random effect variances
)
```



## Monolingual vs Multilingual children: The proportion assessed where the assessment language matches main home language

Here we do the same as above but disaggregate between monolingual and multilingual children. As expected, the percentage of matches is lower among multilingual children than monolingual children. Among multilingual children the overall language match rate is 76% compared to 93% for monolingual children. 

```{r}
#| label: tbl-lang-match-multi
#| tbl-cap: "Monolingual vs Multilingual children: The proportion assessed where the assessment langauge maches the main home language"

df_multilingual <- tb5_2024 %>%
  rowwise() %>%
  mutate(lang_count = sum(c_across(all_of(lang_cols)), na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(multilingual = ifelse(lang_count > 1, 1, 0)) %>%
  select(id_child, multilingual)

# 2) join and compute language match
tb5_2024_lang_all <- tb5_2024_lang %>%
  left_join(df_multilingual, by = "id_child") %>%
  # if some id_child didn't match, assign them monolingual (or NA) as you prefer:
  mutate(multilingual = replace_na(multilingual, NA))

# 3) summarise by province, assessment language and multilingual flag
summary_long <- tb5_2024_lang_all %>%
  mutate(language_match = as.integer(
    as.character(elom_language_assessment) == as.character(cg_language_home)
  )) %>%
  filter(!is.na(cg_language_home), cg_language_home != "", cg_language_home != "None") %>%
  group_by(elom_prov, elom_language_assessment, multilingual) %>%
  summarise(
    prop_language_match = round(100 * mean(language_match, na.rm = TRUE), 1),
    n_language = n(),
    .groups = "drop"
  )

# 4) pivot wider so multilingual=1 and multilingual=0 become separate columns
tbl_wide <- summary_long %>%
  pivot_wider(
    names_from = multilingual,
    values_from = c(prop_language_match, n_language),
    names_glue = "{.value}_{multilingual}",
    values_fill = list(prop_language_match = NA_real_, n_language = 0)
  ) %>%
  rename(
    Province = elom_prov,
    `Assessment language` = elom_language_assessment,
    `Multilingual: % most used home language matches assessment language` = prop_language_match_1,
    `Monolingual: % most used language matches assessment language`     = prop_language_match_0,
    `Multilingual: N` = n_language_1,
    `Monolingual: N`  = n_language_0
  )

# 5) palette and flextable styling (apply to both percent columns)
pal <- col_numeric(
  palette = c("#b2182b", "#fddbc7", "#ffffff"),
  domain  = c(0, 100)
)

# build flextable
ft <- flextable(tbl_wide) %>%
  # format percent columns
  colformat_num(
    j = c("Multilingual: % most used home language matches assessment language",
          "Monolingual: % most used language matches assessment language"),
    digits = 1, suffix = "%"
  ) %>%
  # format N columns as integers
  colformat_num(j = c("Multilingual: N", "Monolingual: N"), digits = 0, na_str = "0")

# Apply background gradient to both percent columns
for (col_name in c("Multilingual: % most used home language matches assessment language",
                   "Monolingual: % most used language matches assessment language")) {
  col_vals <- tbl_wide[[col_name]]
  ft <- bg(
    ft,
    j = col_name,
    i = seq_len(nrow(tbl_wide)),
    bg = pal(col_vals)
  ) %>%
    color(
      j = col_name,
      i = which(!is.na(col_vals) & col_vals <= 25),
      color = "white"
    )
}

ft <- ft %>% autofit()

# ft is the resulting flextable
ft

```


### Monolingual vs multilingual: The impact of language mismatch on ELOM outcomes

Here we add an interaction effect to the earlier mixed-effects model to see whether the impact of a language mismatch differs when the child is monolingual vs multilingual. The hypothesis tested here is that multilingual children would be less affected because they are more likely to be at least familiar with the assessment language compared to monolingual children. Note though that the sample of monolingual children is rather small (10% of PCG sample). 

The model's results now also show that multilingual children, on average, score **better** than monolingual children and that the impact of an assessment language-main home language mismatch **is not** statistically significantly different between monolingual and multilingual children (though the coefficient does tentatively suggest that the affect is greater on monolingual children compared to multilingual children). 

```{r}

tb5x <- tb5_2024 %>%
  left_join(tb5_2024_lang, by = "id_child") %>%
  left_join(df_multilingual, by = "id_child") %>%
  as_factor() %>%
  mutate(
    # language_match (1 = match, 0 = not match)
    language_match = as.integer(
      as.character(elom_language_assessment.y) == as.character(cg_language_home.y)
    ),
    # outcome: 1 = On Track
    ontrack = as.integer(n_elom_total_cuts == "On Track"),
    # tidy covariates
    elom_child_sex = as.factor(elom_child_sex),
    n_pri_fee_group7 = as.factor(n_pri_fee_group7)
  )

# Model with interaction
mod <- lmer(
  n_elom_total_score ~ elom_child_age + elom_child_sex + n_elom_zhfa + n_pri_fee_group7 +
    language_match * multilingual +
    (1 | id_ecd) + (1 | id_ea) + (1 | stratum),
  data = tb5x
  # family = binomial(link = "logit") # if outcome were binary
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e5))
)

# Table output
tab_model(
  mod,
  dv.labels  = "ELOM total score",
  pred.labels = c(
    "(Intercept)",
    "Child age (months)"         = "elom_child_age",
    "Girl (ref: Boy)"            = "elom_child_sexGirls",
    "Height-for-age (z)"         = "n_elom_zhfa",
    "Fee group (factor)"         = "n_pri_fee_group7",
    "Assessment language = home language" = "language_match",
    "Multilingual"               = "multilingual",
    "Language match × Multilingual" = "language_match:multilingual"
  ),
  show.icc   = TRUE,
  show.r2    = TRUE,             
  show.se    = TRUE,
  show.re.var = TRUE
)


```

# The proportion assessed where the assessment language matches ANY home language in the PCG form
The table below shows the % of assessed children for whom the assessment language matches any language the child speaks at home (as per the PCG form, cg_language_home). 

```{r}
#| label: tbl-lang-match-any
#| tbl-cap: "The proportion assessed where the assessment language matches any home language"

tb5_2024_lang <- tb5_2024 %>% 
  select(id_child, elom_prov, cg_language_home, elom_language_assessment) %>% 
  as_factor() 
lang_cols <- paste0("cg_languages_", 1:11)

tb5_2024_home_langs <- tb5_2024 %>%
  select(id_child, all_of(lang_cols)) %>%
  pivot_longer(
    cols = starts_with("cg_languages_"),
    names_to = "name",
    values_to = "value"
  ) %>%
  # robustly coerce value -> integer 0/1
  mutate(value = case_when(
    is.logical(value) ~ as.integer(value),
    is.numeric(value) ~ as.integer(value),
    TRUE               ~ as.integer(as.character(value))
  )) %>%
  filter(!is.na(value) & value == 1) %>%
  group_by(id_child) %>%
  # create an index for each selected language (1,2,3, ...)
  mutate(lang_index = row_number()) %>%
  ungroup() %>%
  # extract numeric code from the column name and map to language labels
  mutate(
    cg_lang_code = as.integer(str_remove(name, "^cg_languages_")),
    cg_language_coerce = factor(
      cg_lang_code,
      levels = 1:11,
      labels = c("Afrikaans", "English", "isiNdebele", "isiXhosa", "isiZulu",
                 "Sesotho", "Sepedi", "Setswana", "siSwati", "Tshivenda", "Xitsonga")
    )
  ) %>%
  select(id_child, lang_index, cg_language_coerce) %>%
  pivot_wider(
    id_cols = id_child,
    names_from = lang_index,
    names_prefix = "cg_language_",
    values_from = cg_language_coerce,
    values_fill = NA_character_
  ) %>% 
  rowwise() %>%
  mutate(cg_language_home_collapse = paste(na.omit(c_across(starts_with("cg_language_"))), collapse = " ")) %>%
  ungroup() %>% 
  select(id_child, cg_language_home_collapse)

tb5_2024_lang <- tb5_2024_lang %>%
  left_join(tb5_2024_home_langs, by = "id_child") %>%
  # mutate(cg_language_coerce = factor(cg_language_coerce,
  #                                    levels = levels(cg_language_home)),
  #        cg_language_home = if_else(is.na(cg_language_home),
  #                                   cg_language_coerce, cg_language_home)) %>%
  # select(-cg_language_coerce) %>% 
  mutate(elom_language_assessment =
           forcats::fct_recode(elom_language_assessment,
                      "Sepedi" = "Sesotho se Leboa (Sepedi)"))


# build the summary table
tbl <- tb5_2024_lang %>% 
  mutate(language_match = as.integer(as.character(elom_language_assessment) %in%
                                     as.character(cg_language_home_collapse))) %>% 
  filter(!is.na(cg_language_home_collapse), cg_language_home_collapse != "", cg_language_home_collapse != "None") %>% 
  group_by(elom_prov, elom_language_assessment) %>% 
  summarise(prop_language_match = round(100 * mean(language_match), 1),
            n_language = n(), .groups = "drop") %>% 
  rename(
    Province = elom_prov,
    `Assessment language` = elom_language_assessment,
    `% most used home language matches assessment language` = prop_language_match,
    N = n_language
  )

# color palette: low values = red, high = white 
pal <- col_numeric(
  palette = c("#b2182b", "#fddbc7", "#ffffff"), # deep red → light red → white
  domain  = c(0, 100)
)

# build flextable
ft <- flextable(tbl) %>%
  colformat_num(j = "% most used home language matches assessment language",
                digits = 1, suffix = "%")

# apply background gradient to that column
col_vals <- tbl$`% most used home language matches assessment language`
ft <- bg(
  ft,
  j = "% most used home language matches assessment language",
  i = seq_len(nrow(tbl)),
  bg = pal(col_vals)
)

# optional: improve contrast for very low values (white text on dark red)
ft <- color(
  ft,
  j = "% most used home language matches assessment language",
  i = which(col_vals <= 25),  # threshold is your call
  color = "white"
) %>%
  autofit()

ft


```